---
title: "Common Text Mining Visualizations"
author: "P. Adames"
date: "June 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set Options
options(stringsAsFactors = FALSE)
Sys.setlocale('LC_ALL', 'C')

library(stringi)
library(stringr)
library(qdap)
library(curl)
library(ggplot2)
library(ggthemes)
library(tm)

source("../Chapter\ 2\ Basics\ of\ Text\ Mining/Helpers.R")
```
## Text Frequency revisited

To solve the exercises from this chapter we will be reusing the code to clean up a corpus of text 
from Chapter 2: Basics of Text Mining. How effective data communicates meaning to any given 
audiece is a strong function of the skills and habits when it comes to consuming and interpreting information. This may have a decisive impact on the decisions made based on the data.

To give a concrete example of this, we will contrast the table of top 20 most frequent terms in the 
corpus of Delta Airlines Customer Service Tweets presented in the previous chapter with a plot of the same data.

```{r summary of cleaup code and term frequency data frame}
text.df <- read.csv(curl("https://raw.githubusercontent.com/padames/text_mining/master/oct_delta.csv"),
                    header = T, stringsAsFactors = F)
df <- data.frame(doc_id = seq(1:nrow(text.df)), # a numeric index
                 text = text.df$text) # the vector of 1377 tweets
corpus <- VCorpus(DataframeSource(df))
corpus.clean <- clean.corpus(corpus)
tdm <- TermDocumentMatrix(corpus.clean)
tdm.m <- as.matrix(tdm)
term.freq <- rowSums(tdm.m)
frequencies.df <- data.frame( word = names(term.freq), frequency = term.freq)
frequencies.df <- frequencies.df[order(term.freq, decreasing = T), ]
```

Now we can write the new code to build the `ggplot` visualization:

```{r ggplot visualization}
frequencies.df$word <- factor(x = frequencies.df$word, 
                              levels = unique(as.character(frequencies.df$word)))
ggplot(frequencies.df[1:20,], aes(x = word, y = frequency)) +
  geom_col(stat = "identity", colour = "darkred") +
  coord_flip() +
  theme_gdocs() +
  geom_text(aes(label = frequency), colour = "white", hjust = 1.25, size = 5.0 )
```
